Bricking up backdoors in Intrusion Detection Systems

### 2019-09-10 ex2/paper/paper.tex
Make it fit to 6 pages

%For research,
The invention and assessment of techniques for \glspl{ids} poses many challenges. Creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power. In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into another \gls{ad} software. % which might even be open-source.
If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have implemented backdoors which circumvent the \gls{ad} system. This could be motivated by profitseeking or by government actors requiring ways to purposefully disable security measures in specific cases.

@@

%\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the CIC-IDS-2017 classifier.} % The top plot shows the full range of TTL values while for the lower plot only \gls{ttl} values from 0 to 5 are displayed.}
\label{fig:pdp_ttl}
\end{figure}
Backdoors can be identified by computing \gls{pdp} or \gls{ale} plots for the \gls{mui} and investigating if regions exist, for which the \gls{mui} behaves counter-intuitive. For our CIC-IDS-2017 \gls{dl} classifier, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction. We also provide plots for the corresponding models which were trained without backdoor. The plots are not available in a real situation, but we provide them here for comparison.

@@

\begin{figure}[b]
\includegraphics[width=\columnwidth]{figures/ttlmean.pdf}
\caption{\gls{pdp} and \gls{ale} plots for mean(TTL) for the CIC-IDS-2017 \gls{rf} and \gls{dl} classifiers.}
\label{fig:ttlmean}
\end{figure}
Independent of the origin of such patterns, they might be exploited for masquerading attacks and thus are clearly unwanted. The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models for such patterns.
\subsection{Pruning Techniques}

To perform pruning, a validation dataset is needed. We take a validation dataset that is one fourth the size of the training dataset. We use the validation dataset for pruning as described in the next sections and use a test dataset that is also one fourth to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers. Training, validation and test dataset are pairwise disjoint.

@@

%We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above. 
The concept behind our pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease
%by a lot
significantly,
while decisions that are used for the backdoor are likely to be removed.
We developed several variants of our pruning defense with increasing levels of sophistication:
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item Like (1) but consider only ``harmless'' leaves: We assume that attackers want malicious samples appear harmless.
\item Like (2) but use depth to decide when two leaves used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and are thus have lower depth.
\end{enumerate}
\begin{figure*}[h]
\includegraphics[width=\textwidth]{pruning_example.pdf}
\caption{Example of a \gls{dt} being pruned. To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf. It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside). The leaf that is going to be cut in each step has a dashed connection to its parent.}
\label{fig:pruningExample}
\end{figure*}

\autoref{fig:pruningExample} shows an example of variant (3) of pruning applied to a tree
\begin{figure}[h]
\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
\caption{Pruning process an \gls{rf} for \unsw{}, considering only harmless leaves and depth. Dash length proportional to validtion set size (from 0.01 until 1).}
\label{fig:performancePruningRF}
\end{figure}
%\todo{Max: Which pruning algorithm actually?}
and \autoref{fig:performancePruningRF} shows the resulting accuracy for pruning according to variant (3) for the \unsw{} dataset. With sufficient neurons being pruned, the accuracy of the backdoor approaches zero. The accuracy for the regular data does not decrease, it even increases by
%thanks to the the pruning also
reducing overfitting. %With a significantly smaller validation set, pruning takes longer to remove the backdoor
%until it works
%but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
Even with a very small validation dataset of just 1\% of its original size, the pruning still works as expected.

@@

>>>>

%For research,
%The invention and assessment of techniques for \glspl{ids} poses many challenges.
Training a machine learning model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power. In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into another \gls{ad} software. % which might even be open-source.
If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have implemented backdoors which circumvent the \gls{ad} system. This could be motivated by profitseeking or by government actors requiring ways to purposefully disable security measures in specific cases.

@@

%\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the \cic{}. Full range of TTL values on top; only TTL values from 0 to 5 below.}
\label{fig:pdp_ttl}
\end{figure}
Backdoors can be identified by computing \gls{pdp} or \gls{ale} plots for the \gls{mui} and investigating if regions exist, for which the \gls{mui} behaves counter-intuitive. For our CIC-IDS-2017 \gls{dl} classifier, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction. We also provide plots for the corresponding models which were trained without backdoor. The plots are not available in a real situation, but we provide them here for comparison.

@@

\begin{figure}[b]
\includegraphics[width=\columnwidth]{figures/ttlmean.pdf}
\caption{\gls{pdp} and \gls{ale} plots for mean(TTL) for the CIC-IDS-2017 \gls{rf} and \gls{dl} classifiers.}
\label{fig:ttlmean}
\end{figure}
Independent of the origin of such patterns, they might be exploited for masquerading attacks and thus are clearly unwanted. The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models for such patterns.

\begin{figure*}[h]
\includegraphics[width=\textwidth]{pruning_example.pdf}
\caption{Example of a \gls{dt} being pruned. To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf. It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside). The leaf that is going to be cut in each step has a dashed connection to its parent.}
\label{fig:pruningExample}
\end{figure*}

\subsection{Pruning Techniques}
To perform pruning, a validation dataset is needed. We take a validation dataset that is one fourth the size of the training dataset. We use the validation dataset for pruning as described in the next sections and use a test dataset that is also one fourth to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers. Training, validation and test dataset are pairwise disjoint.

@@

%We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above. 
The concept behind our pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease
%by a lot
significantly,
while decisions that are used for the backdoor are likely to be removed.
We developed several variants of our pruning defense with increasing levels of sophistication:
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item Like (1) but consider only ``harmless'' leaves: We assume that attackers want malicious samples appear harmless.
\item Like (2) but use depth to decide when two leaves used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and are thus have lower depth.
\end{enumerate}

\autoref{fig:pruningExample} shows an example of variant (3) of pruning applied to a tree
\begin{figure}[h]
\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
\caption{Pruning an \gls{rf} for \unsw{}.}
\label{fig:performancePruningRF}
\end{figure}
%\todo{Max: Which pruning algorithm actually?}
and \autoref{fig:performancePruningRF} shows the resulting accuracy for pruning according to variant (3) for the \unsw{} dataset. With sufficient neurons being pruned, the accuracy of the backdoor approaches zero. The accuracy for the regular data does not decrease, it even increases by
%thanks to the the pruning also
reducing overfitting. %With a significantly smaller validation set, pruning takes longer to remove the backdoor
%until it works
%but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
Even with a very small validation dataset of just 1\% of its original size, the pruning still works as expected.

@@

The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure. Hence, a decay of the backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
